---
title: "Draw a single triplet"
author: "Ignatius Pang"
date: "9 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCy3)

```


## Reference
I've used the Bioconductor RCy3 vignette extensively for this demo: https://bioconductor.org/packages/release/bioc/vignettes/RCy3/inst/doc/RCy3.R

## Add the nodes
```{r }
g <- new('graphNEL', edgemode='undirected')

g <- graph::addNode( 'A', g)
g <- graph::addNode( 'B', g)
g <- graph::addNode( 'C', g)

# Create a Cytoscape Window
cw <- CytoscapeWindow('One Triplet', graph=g, overwrite=TRUE)

# Push the network into to Cytoscape 
displayGraph(cw)

# Use layout algorithm to draw the graph
layoutNetwork(cw, layout.name = "force-directed")

```

## Add node attributes
```{r }

g <- cw@graph   # created above, in the section  'A minimal example'

g <- initNodeAttribute (graph=g,  attribute.name='Gene_Symbol',
                            attribute.type='char',
                            default.value='undefined')

g <- initNodeAttribute (graph=g,  attribute.name='moleculeType',
                            attribute.type='char',
                            default.value='undefined')

g <- initNodeAttribute (graph=g,  'log_fold_change', 'numeric', 0.0)

#  oln_id_a | oln_id_b | oln_id_c | type_ac | type_bc |  score   
# ----------+----------+----------+---------+---------+---------
#  YKL101W  | YDL056W  | YDR507C  | kd      | td      | -0.3311


# Add the gene name 
nodeData (g, 'A', 'Gene_Symbol') <- 'HSL1' # YKL101W
nodeData (g, 'B', 'Gene_Symbol') <- 'MBP1' # YDL056W
nodeData (g, 'C', 'Gene_Symbol') <- 'GIN4' # YDR507C

# Add molecule type
nodeData (g, 'A', 'moleculeType') <- 'kinase'
nodeData (g, 'B', 'moleculeType') <- 'TF'
nodeData (g, 'C', 'moleculeType') <- 'protein'

# Add fold change
nodeData (g, 'A', 'log_fold_change') <- -1.4
nodeData (g, 'B', 'log_fold_change') <- 1.9
nodeData (g, 'C', 'log_fold_change') <- 3.7

# Send to Cytoscape
cw <- setGraph (cw, g)
displayGraph (cw)    # cw's graph is sent to Cytoscape

layoutNetwork(cw, layout.name = "force-directed")

```

# Modifying Node Display Style
```{r}


# Set defaults
    setDefaultNodeShape (cw, 'ELLIPSE')
    setDefaultNodeColor (cw, '#AAFF88')
    setDefaultNodeSize  (cw, 80)
    setDefaultNodeFontSize (cw, 40)

# Change shape of node
    getNodeShapes (cw)   # diamond, ellipse, trapezoid, triangle, etc.
    print (noa.names (getGraph (cw)))  # what data attributes are defined?
    print (noa (getGraph (cw), 'moleculeType')) # Print the node attributes for each node
    attribute.values <- c ('kinase',  'TF',       'protein')
    node.shapes      <- c ('DIAMOND', 'TRIANGLE', 'CIRCLE')
    setNodeShapeRule (cw, node.attribute.name='moleculeType',
                      attribute.values, node.shapes)

    
# Change the colour of the node according to the log fold-change in transcript abundances
    setNodeColorRule (cw, 'log_fold_change', c (-3.0, 0.0, 3.0),
                      c ('#00AA00', '#00FF00', '#FFFFFF', '#FF0000', '#AA0000'),
                      mode='interpolate')
    
# You can also change the size of the node based on the log fold-change value    
    # control.points = c (-1.2, 2.0, 4.0)
    # node.sizes     = c (10, 20, 50, 200, 205)
    # setNodeSizeRule (cw, 'log_fold_change', control.points, node.sizes,
    #                  mode='interpolate')

```


## Add edges and edge attributes
```{r}


    g <- cw@graph
    g <- initEdgeAttribute (graph=g,  attribute.name='edgeType',
                            attribute.type='char',
                            default.value='unspecified')
    
    # Add edges
    g <- graph::addEdge ('A', 'B', g)
    g <- graph::addEdge ('B', 'C', g)
    g <- graph::addEdge ('A', 'C', g)

    # Assign edge attributes
    edgeData (g, 'A', 'B', 'edgeType') <- 'negative genetic interaction'
    edgeData (g, 'B', 'C', 'edgeType') <- 'activates transcription of target gene'
    edgeData (g, 'A', 'C', 'edgeType') <- 'activates via phosphorylation'
    
    
    cw@graph <- g
    displayGraph (cw)
```

## Set the Style of the lines
```{r}
    # 
    getLineStyles(cw)
    
    line.styles <- c ( 'SOLID', 'DOT',  'PARALLEL_LINES')
    edgeType.values <- c ('activates via phosphorylation', 'activates transcription of target gene',
                         'negative genetic interaction')
    
    setEdgeLineStyleRule (cw, 'edgeType', edgeType.values,
                          line.styles)

    getArrowShapes(cw)
    
    ## Source Arrow Styles
    source.arrow.styles <- c ('None', 'None', 'None') # c ('NONE', 'NONE', 'NONE')

    setEdgeSourceArrowRule (cw, 'edgeType', 
    						attribute.values=edgeType.values,
                            arrows=source.arrow.styles, 
    						default="NONE")
    
    ## Target Arrow Styles
    target.arrow.styles <- c ('ARROW', 'DELTA', 'NONE')
    setEdgeTargetArrowRule (cw, 'edgeType', 
    						edgeType.values,
                            target.arrow.styles, 
    						default="NONE")

    cw@graph <- g
    displayGraph (cw)
   
    layoutNetwork(cw, layout.name = "force-directed")
    
```



### animation, change the position of the nodes
```{r}

    cwe <- cw

    ## You can build a simple graph but I've commented it out heres
    #cwe <- CytoscapeWindow ('vignette.setNodePosition',
    #                        graph=RCy3::makeSimpleGraph (), overwrite=TRUE)
    #displayGraph (cwe)
    #layoutNetwork (cwe, 'grid')

    center.x <- 100
    center.y <- 100
    radius <- 100
    # sweep through full revoltion 3 times, 5 degrees at a time
    angles <- seq (0, 360, 10)
    for (angle in angles) {
        angle.in.radians <- angle * pi / 180
        x <- center.x + (radius * cos (angle.in.radians))
        y <- center.y + (radius * sin (angle.in.radians))
        setNodePosition (cwe, 'A', x, y)
        system ('sleep 0.3')

    }
    # RCy3 will not create windows with duplicate names, so clear the decks for a subsequent possible run
```


###  moviePrep (eval = FALSE)
```{r}
    cwm <- cw

    # g <- RCy3::makeSimpleGraph ()
    # g <- initNodeAttribute (g, 'pval', 'numeric', 1.0)
    # 
    # cwm <- CytoscapeWindow ('movie', graph=g, overwrite=TRUE)
    # displayGraph (cwm)
    # layoutNetwork (cwm, 'grid')
```


### movieRules: Rules for changing the color and node size
```{r}
    cwm <- cw

    lfc.control.points <- c (-3.0, 0.0, 3.0)
    lfc.colors <- c ('#00AA00', '#00FF00', '#FFFFFF', '#FF0000', '#AA0000')
    setNodeColorRule (cwm, 'log_fold_change', lfc.control.points, lfc.colors,
                      mode='interpolate')

    pval.control.points <- c (0.1, 0.05, 0.01, 0.0001)
    pval.sizes          <- c (30, 50, 70, 100)
    setNodeSizeRule (cwm, 'pval', pval.control.points, pval.sizes,
                     mode='interpolate')
```

### animate, change color and node size
```{r}

    pval.timepoint.1 <- c (0.01, 0.3, 0.05)
    pval.timepoint.2 <- c (0.05, 0.01, 0.01)
    pval.timepoint.3 <- c (0.0001, 0.005, 0.1)

    lfc.timepoint.1 <- c (-1.0, 1.0, 0.0)
    lfc.timepoint.2 <- c (2.0, 3.0, -2.0)
    lfc.timepoint.3 <- c (2.5, 2.0, 0.0)

    for (run in 1:10 ) { 
    for (i in 1:5) {  # run this loop 5 times
        setNodeAttributesDirect (cwm, 'log_fold_change',  'numeric', c ('A', 'B', 'C'),
                                 lfc.timepoint.1)
        setNodeAttributesDirect (cwm, 'pval', 'numeric', c ('A', 'B', 'C'),
                                 pval.timepoint.1)
        system ('sleep 0.3')

        setNodeAttributesDirect (cwm, 'log_fold_change',  'numeric', c ('A', 'B', 'C'),
                                 lfc.timepoint.2)
        setNodeAttributesDirect (cwm, 'pval', 'numeric', c ('A', 'B', 'C'),
                                 pval.timepoint.2)
        system ('sleep 0.3')

        setNodeAttributesDirect (cwm, 'log_fold_change',  'numeric', c ('A', 'B', 'C'),
                                 lfc.timepoint.3)
        setNodeAttributesDirect (cwm, 'pval', 'numeric', c ('A', 'B', 'C'),
                                 pval.timepoint.3)
        system ('sleep 0.3')
    }
    }

```

## Delete an edge
```{r}
selectEdges (cw, "A (negative genetic interaction) B")
deleteSelectedEdges (cw)

```

## Get the selected node
```{r}
getSelectedNodes(cw)
```


## Get the selected Node and find the first neighbour, then delete it 
```{r}
selectFirstNeighborsOfSelectedNodes(cw)

deleteSelectedNodes (cw)
```

### code chunk number 19: deleteWindows (eval = FALSE)
```{r}
    cy <- CytoscapeConnection ()
    window.names <- c ('One Triplet')
    for (window.name in window.names){
        if (window.name %in% as.character (getWindowList (cy))){
            deleteWindow (cy, window.name)
        }
    }
```
