---
title: "igraph example"
author: Ignatius Pang
output: html_document

---

Install and Load Igraph
```{r}
# Install the igraph library
# install.packages('igraph')

# load the igraph library
library(igraph)
```

Configure directories 
```{r}
# Configure the base directory 
setwd("/home/ignatius/PostDoc/2017/Network_SURF")

base_directory <- getwd()
 
# Configure the data directory
data_directory <- file.path( base_directory, "Data/SURF/igraph" ) 

# Configure the results directory
results_directory <- file.path( base_directory, "Results/SURF/igraph" ) 

# Create the results directory if it does not yet exists
if(      ! dir.exists(results_directory) ) {
		dir.create(results_directory, recursive = TRUE)
}
```

Read the input files with the node and edge information 
```{r}
# Read the Nodes data
people <- read.csv( file.path( data_directory, "people.csv"), head=TRUE)

# Read the Edges data
relationships <- read.csv(file.path(data_directory, "relationships.csv") , head=TRUE)

# check what the tables look like

head(people)
head(relationships)
```


```{r}
#### Create an empty graph object
g <- graph.empty(directed=FALSE)

### Add vertices
g <- add.vertices(g, nrow(people),
                     name=as.character(people[,1]), age=people[,2],
                     gender=as.character(people[,3]))



### Add edges
  edges <- t(relationships[,c('Person_A', 'Person_B')])


 g <- add.edges(g, edges,
               social.group=as.character(relationships[,3]),
	       num.emails=as.character(relationships[,4]))


 # View nodes and edges data
 print ( "Look at the node properties")
 print ("Get the List of Names")
 V(g)$name
 
 print ( "Get the gender of the people")
 V(g)$gender
 
 print ( "Get the age of the people")
 V(g)$age
 
 # Look at Edge Properties 
 print ( "Look at the Edge Properties")
 print ( "Get the relationship between two people")
 E(g)$social.group
 
 print ( "Get the number of e-mails sent between two people ")
 E(g)$num.emails
 
 # Plot the graph (draft)
 plot(g)
```

```{r}
# Add a node and an edge 

g <- g + vertex("Kylie", gender="F", age=23) +
	     edge("David", "Kylie", num.emails=2, social.group="Client")
	
plot(g)


```

```{r}
# Remove an edge: Helen--Iris

 g_temp <- g

 g <- g -edge("Helen|Iris")

 plot(g)

# Other ways of removing an edge 
 g_1 <- g_temp - path ( "Helen", "Iris") 

 g_2 <- g_temp %>% delete_edges("Helen|Iris")  

 g_3 <- g_temp -edge("Iris|Helen")
 
 
 # Remove the node named Iris 
 
 g <- g - vertex('Iris')
	
plot(g)


```


We want to format the graph based on the node properties 
```{r}
# Colour nodes by gender
V(g)$color <- ifelse(V(g)$gender == 'M', "cyan", "pink")

# Node size represents age
V(g)$size <- V(g)$age

# Plot the graph (draft)
plot(g)
```

We want to format the graph based on the edge properties 
```{r}
# Edge thickness represents number of e-mails
E(g)$width <- E(g)$num.emails

# Edge colour represents social group
edge_color_lookup <- list( tennis ="orange", basketball="blue", client="black", fencing="brown" )
E(g)$color <- unlist(lapply(E(g)$social.group, function(x){ edge_color_lookup[[tolower(x)]] } ))

plot(g)
```



Some network statistics 
```{r}
library(igraph)
# Calculate the degree and betweenness scores of nodes in the network

# Number of friends for each person
degree(g)

# Centrality measure of each person. Betweeness is the number of shortest path that pass throught this node. It seems that David is central to this social network. 
betweenness(g)
```

Save the results in output directories
```{r}
# Save the graph visualization in a PDF file
pdf(paste(results_directory, "example.pdf", sep="") )
dev.off()

## This file can be opened in Cytoscape
#http://opentutorials.cgl.ucsf.edu/index.php/Portal:Cytoscape3
write.graph(g, paste(results_directory, "example.graphml", sep="") , format= "graphml")

# Get the adjacency matrix 
adjacency_matrix <- get.adjacency(g) %>% as.matrix()
adjacency_matrix

# Get the edge list
edge_list <- get.edgelist(g)
edge_list

```


# Based on the as_graphnel object from igraph package.
# Edited such that it deals with class attribute 
```{r}
my_as_graphnel <- function (graph) 
{
    if (!is_igraph(graph)) {
        stop("Not an igraph graph")
    }
    if ("name" %in% vertex_attr_names(graph) && is.character(V(graph)$name)) {
        name <- V(graph)$name
    }
    else {
        name <- as.character(seq(vcount(graph)))
    }
    edgemode <- if (is_directed(graph)) 
        "directed"
    else "undirected"
    if ("weight" %in% edge_attr_names(graph) && is.numeric(E(graph)$weight)) {
        al <- lapply(as_adj_edge_list(graph, "out"), as.vector)
        for (i in seq(along = al)) {
            edges <- ends(graph, al[[i]], names = FALSE)
            edges <- ifelse(edges[, 2] == i, edges[, 1], edges[, 
                2])
            weights <- E(graph)$weight[al[[i]]]
            al[[i]] <- list(edges = edges, weights = weights)
        }
    }
    else {
        al <- as_adj_list(graph, "out")
        al <- lapply(al, function(x) list(edges = as.vector(x)))
    }
    names(al) <- name
    res <- new("graphNEL", nodes = name, edgeL = al, edgemode = edgemode)
    g.n <- graph_attr_names(graph)
    if ("directed" %in% g.n) {
        warning("Cannot add graph attribute `directed'")
        g.n <- g.n[g.n != "directed"]
    }
    for (n in g.n) {
        res@graphData[[n]] <- graph_attr(graph, n)
    }
    v.n <- vertex_attr_names(graph)
    v.n <- v.n[v.n != "name"]
    for (n in v.n) {

    	# graph::nodeDataDefaults(res, attr = n) <- NA
		#attr(graph::nodeDataDefaults(res, attr = n), 'class') <- 'char'

        test_node_data_type <- vertex_attr(graph, n)[1]
        	        
			        if ( is.numeric(test_node_data_type)) {
			        	graph::nodeDataDefaults(res, attr = n) <- 0
			        	attr(graph::nodeDataDefaults(res, attr = n), 'class') <- 'numeric'

			        } else if (is.integer(test_node_data_type)) {
			        	graph::nodeDataDefaults(res, attr = n) <- 0 
			        	attr(graph::nodeDataDefaults(res, attr = n), 'class') <- 'integer'

			        } else { # if ( is.character(test_node_data_type)) {
			        	graph::nodeDataDefaults(res, attr = n) <- ''
			        	attr(graph::nodeDataDefaults(res, attr = n), 'class') <- 'char'
					}

        print ( vertex_attr(graph, n))
        
        graph::nodeData(res, attr = n) <- vertex_attr(graph, n)
    }
    e.n <- edge_attr_names(graph)
    e.n <- e.n[e.n != "weight"]
    if (length(e.n) > 0) {
        el <- as_edgelist(graph)
        el <- paste(sep = "|", el[, 1], el[, 2])
        
        # Initialize edge weight attribute anyway as it is always used in cytoscape
         graph::edgeDataDefaults(res, attr="weight") <- 0
         attr(graph::edgeDataDefaults(res, attr="weight"), 'class') <- 'numeric'
        
        for (n in e.n) {

        	  # graph::edgeDataDefaults(res, attr = n) <- NA
        	#	attr(graph::edgeDataDefaults(res, attr = n), 'class') <- 'char'

        	   test_edge_data_type <- edge_attr(graph, n)[1]

        	        
				        if ( is.numeric(test_edge_data_type)) {
				        	graph::edgeDataDefaults(res, attr = n) <- 0
				        	attr(graph::edgeDataDefaults(res, attr = n), 'class') <- 'numeric'
				        } else if (is.integer(test_edge_data_type)) {

 					        graph::dgeDataDefaults(res, attr = n) <- 0
				        	attr(graph::edgeDataDefaults(res, attr = n), 'class') <- 'integer'
 					        

				        } else { # if ( is.character(test_edge_data_type)) {
				        	graph::edgeDataDefaults(res, attr = n) <- ''
				        	attr(graph::edgeDataDefaults(res, attr = n), 'class') <- 'char'
				        	
				        }
        
        		
			print ( edge_attr(graph, n))
			
            res@edgeData@data[el] <- mapply(function(x, y) {
                xx <- c(x, y)
                names(xx)[length(xx)] <- n
                xx
            }, res@edgeData@data[el], edge_attr(graph, n), SIMPLIFY = FALSE)
        }
    }
    res
}
```

```{r}
library(RCy3)

GNEL2 <-my_as_graphnel(g)

# Convert graphNEL object back to igraph object and I can see all attributes are still preserved
g4 <- graph_from_graphnel(GNEL2)
plot(g4)

#nodeData(GNEL2)
#str(GNEL2)

# Move it into a Cytoscape Window
new_empth_graph <- new('graphNEL', edgemode='undirected')

cw <- CytoscapeWindow ( 'Mini Social Network', graph=new_empth_graph, overwrite=TRUE)
cw <- setGraph (cw, GNEL2)
cw@graph <- GNEL2
RCy3::displayGraph(cw)

RCy3::layoutNetwork(cw, layout.name = "force-directed")


```



